Fitting metabolic models to dissolved oxygen data: The Estuarine BAyesian Single-station Estimation method (EBASE)

### 2024-03-12 revision/review_response.qmd

*Response: Apologies for this confusion. Eq. 3 is not used in EBASE, rather it is discussed as the gas transfer used in BASE that is inappropriate for our application, justifying use of the gas exchange model in @Wanninkhof14. Line 166 was revised to make this more clear: "Based on this difference and the lack of availability of other possible drivers of the transfer velocity in coastal systems, Equation 3 is not used herein. Instead, the well-established wind- and temperature-based parameterization for gas exchange of @Wanninkhof14 was used for EBASE (Equation 2)..."*

Diving into eq 3, higher water levels usually indicate higher flows and thus higher gas transfer velocity (Kw). I’m not exactly sure how temperature plays into this equation, and if it is appropriate for an estuary. It seems to suggest warm periods within a day have higher exchange rates, due to differences in kinetics at the water surface. However warmer temperatures may also reflect increased stratification and less vertical mixing (small lakes). There may be something to clean from changes in temperature that drive changes in turbulence, but simply using warmer = more exchange seems like an oversimplification. Eq 3 also includes a term related to water depth (H). In flowing waters this equation suggests that Kw increases with water level, likely a result of high-water levels = higher velocities = higher turbulence = higher kw. In an estuary, I’m not sure this relationship truly holds as water depth is not as good of a surrogate for velocity. Highest water level may be at slack tide and thus minimal water movements. It could be that change in water depth (assuming the tidal currents are strong relative to the river inputs) would be a better term to reflect instantaneous discharge. Or possibly flow could be included as a possible input parameter. But I realize this term is not typically available.

*Response: We have noted lack of clarity in our description of Equation 3 that includes the depth-dependency.  As noted by Reviewer 3, @Grace15 do not include a depth-dependency in this equation. We have since revised the equation and text accordingly, emphasizing the importance of wind for the transfer velocity in open-water systems.  Regarding temperature, we feel justified in its inclusion for gas exchange given the previously noted relationship with the Schmidt number, a known factor to be important in open-water systems.  Also see our response to your comment on line 164.*

@@

L143: what is b? and for what time interval is b constant? This ends up being an output (Figure 3) and it is confusing to say that it is constant. So maybe it is constant within a day? Is this a standard parameter for gas exchange? As with the other parameters, providing some type of ‘name’ would be helpful, especially if provided in a table.

*Response: Please see our response to the previous comment and the new input/output table. It's an ambiguous parameter, but really just an indication of how sensitive the gas transfer is to wind speed. The sentence was revised as "The $b$ parameter is fairly well known for open-ocean conditions primarily as an expression of the sensitivity of gas exchange to wind..."*

L164: The equation implies that Kw would increase as water level (H) increases. It seems the authors have this statement inversed or talking about measured reaeration (K). I believe this relates to streams where increasing water levels relate to higher water velocities, and thus higher turbulence and higher gas transfer velocity (kw). See General Comments.

*Response: This was an inaccurate statement on our part.  We have updated the equation and text accordingly:*

*Line 141: "Equation 1 is similar to the oxygen mass balance equation in BASE [@Grace15], with the main difference being the treatment of gas exchange. While both assume that gas exchange is proportional to the concentration gradient across the air–water interface, the treatments of the proportionality differ.  @Grace15 used the following for the volumetric gas flux ($Dvol$) in BASE:*

$$
Dvol = [\, -K \cdot 1.024^{\left(T - \bar{T} \right)}\left(C_{sat} - C \right) ]\,
$$

*where $T$ is the instantaneous water temperature, $\bar{T}$ is the average water temperature over each 24-hour period in the oxygen time series (both in $^{\circ}$C), $C$ is the measured dissolved oxygen, and $C_{sat}$ is the saturation concentration. $K$ is the reaeration coefficient, which was estimated from the data.  In EBASE, gas flux per unit surface area of the air-water interface is proportional to the concentration gradient, with the proportionality taken from the well-established wind- and temperature-based parameterization for gas exchange of @Wanninkhof14 (see Equation 2).  EBASE also allows $b$ to be a free parameter that may reflect spatial and temporal variability in factors other than wind and temperature that may influence transfer velocity. @Wanninkhof14 estimated a fixed $b$ parameter as 0.251 $\left(\text{cm}~\text{hr}^{-1}\right)/\left(\text{m}^2~\text{s}^{-2}\right)$."*

### 2024-03-12 manu-draft.qmd
MH comments added

@eq-ebase is similar to the oxygen mass balance equation in BASE [@Grace15], with the main difference being the treatment of gas exchange. The gas transfer velocity, $k_w$ (@eq-kw), increases as the water column in the vicinity of the air--water interface becomes more turbulent. As such, $k_w$ is usually modeled as a function of measurable variables that can potentially predict turbulent mixing of the surface water, such as wind speed. $k_w$ also increases with temperature because the film of water at the air--water interface becomes thinner as the viscosity decreases and the random motion of gas molecules increases. This dependence of $k_w$ on temperature is usually modeled with the Schmidt number as in @eq-ebase, which combines information on the viscosity of the fluid and the diffusivity of the gas.

@Grace15 used the following for the gas transfer in BASE:

$$
k_w = K \cdot 1.024^{\left(T - \bar{T}\right)}
$$ {#eq-gasexbase}

where $T$ is the instantaneous water temperature, $\bar{T}$ is the average water temperature over each 24-hour period in the oxygen time series (both in $^{\circ}$C), and $K$ is the reaeration coefficient estimated from the data. @eq-gasexbase indicates that $k_w$ increases with temperature, consistent with the known decrease of $k_w$ with the Schmidt number. However, open-water systems like estuaries are more susceptible to wind-based influences on gas exchange, particularly for $k_w$ [@Wanninkhof14; @vandam2019]. Based on this difference and the lack of availability of other possible drivers of the transfer velocity in coastal systems, @eq-gasexbase is not used herein. Instead, the well-established wind- and temperature-based parameterization for gas exchange of @Wanninkhof14 was used for EBASE [@eq-kw], but allowing $b$, which was estimated by @Wanninkhof14 to be 0.251 $\left(\text{cm}~\text{hr}^{-1}\right)/\left(\text{m}^2~\text{s}^{-2}\right)$, to be a free parameter that may reflect spatial and temporal variability in factors other than wind and temperature that may influence transfer velocity.

>>>>

@eq-ebase is similar to the oxygen mass balance equation in BASE [@Grace15], with the main difference being the treatment of gas exchange. While both assume that gas exchange is proportional to the concentration gradient across the air–water interface, the treatments of the proportionality differ. @Grace15 used the following for the volumetric gas flux ($Dvol$) in BASE:

$$
Dvol = [\, -K \cdot 1.024^{\left(T - \bar{T} \right)}\left(C_{sat} - C \right) ]\,
$$ {#eq-gasexbase}

where $T$ is the instantaneous water temperature, $\bar{T}$ is the average water temperature over each 24-hour period in the oxygen time series (both in $^{\circ}$C), $C$ is the measured dissolved oxygen, and $C_{sat}$ is the saturation concentration. $K$ is the reaeration coefficient, which was estimated from the data. In EBASE, gas flux per unit surface area of the air-water interface is proportional to the concentration gradient, with the proportionality taken from the well-established wind- and temperature-based parameterization for gas exchange of @Wanninkhof14 (see @eq-kw). EBASE also allows $b$ to be a free parameter that may reflect spatial and temporal variability in factors other than wind and temperature that may influence transfer velocity. @Wanninkhof14 estimated a fixed $b$ parameter as 0.251 $\left(\text{cm}~\text{hr}^{-1}\right)/\left(\text{m}^2~\text{s}^{-2}\right)$.
